---
title: "Study Case 1"
author: "Mulia"
date: "2022-10-26"
output: html_document
---
### Divvy_Exercise_Full_Year_Analysis ###

# This analysis is based on the Divvy case study "'Sophisticated, Clear, and Polished’: Divvy and Data Visualization" written by Kevin Hartman (found here: https://artscience.blog/home/divvy-dataviz-case-study). The purpose of this script is to consolidate downloaded Divvy data into a single dataframe and then conduct simple analysis to help answer the key question: “In what ways do members and casual riders use Divvy bikes differently?”


# # # # # # # # # # # # # # # # # # # # # # # 
# Install required packages
# tidyverse for data import and wrangling
# lubridate for date functions
# ggplot for visualization
# # # # # # # # # # # # # # # # # # # # # # #  

```{r}
library(tidyverse)  #helps wrangle data
library(lubridate)  #helps wrangle date attributes
library(ggplot2)  #helps visualize data
library("gridExtra")
library("geosphere")
```


```{r}
getwd() #displays your working directory
```
#=====================
# STEP 1: COLLECT DATA
#=====================

```{r}
m10_2021 <- read_csv("202110-divvy-tripdata.csv")
m11_2021 <- read_csv("202111-divvy-tripdata.csv")
m12_2021 <- read_csv("202112-divvy-tripdata.csv")
m1_2022 <- read_csv("202201-divvy-tripdata.csv")
m2_2022 <- read_csv("202202-divvy-tripdata.csv")
m3_2022 <- read_csv("202203-divvy-tripdata.csv")
m4_2022 <- read_csv("202204-divvy-tripdata.csv")
m5_2022 <- read_csv("202205-divvy-tripdata.csv")
m6_2022 <- read_csv("202206-divvy-tripdata.csv")
m7_2022 <- read_csv("202207-divvy-tripdata.csv")
m8_2022 <- read_csv("202208-divvy-tripdata.csv")
m9_2022 <- read_csv("202209-divvy-tripdata.csv")
```

#====================================================
# STEP 2: WRANGLE DATA AND COMBINE INTO A SINGLE FILE
#====================================================
# Compare column names each of the files
# While the names don't have to be in the same order, they DO need to match perfectly before we can use a command to join them into one file

```{r}
colnames(m10_2021)
colnames(m11_2021)
colnames(m12_2021)
colnames(m1_2022)
```
# Inspect the dataframes and look for incongruencies
```{r}
str(m10_2021)
str(m11_2021)
str(m12_2021)
str(m1_2022)
str(m2_2022)
str(m3_2022)
str(m4_2022)
str(m5_2022)
str(m6_2022)
str(m7_2022)
str(m8_2022)
str(m9_2022)
```
# Stack individual quarter's data frames into one big data frame

```{r}
all_trips <- bind_rows(m10_2021, m11_2021, m12_2021, m1_2022, m2_2022, m3_2022, m4_2022, m5_2022, m6_2022, m7_2022, m8_2022, m9_2022)
```

#======================================================
# STEP 3: CLEAN UP AND ADD DATA TO PREPARE FOR ANALYSIS
#======================================================

# Inspect the new table that has been created
```{r}
colnames(all_trips)  #List of column names
```
```{r}
nrow(all_trips)  #How many rows are in data frame?
```
```{r}
dim(all_trips)  #Dimensions of the data frame?
```
```{r}
head(all_trips)  #See the first 6 rows of data frame
```

```{r}
tail(all_trips)
```

```{r}
str(all_trips)  #See list of columns and data types (numeric, character, etc)
```
```{r}
summary(all_trips)  #Statistical summary of data. Mainly for numerics
```
```{r}
#all_trips <- all_trips %>%
#  mutate(member_casual = recode(member_casual
 #                          ,"Subscriber" = "member"
  #                         ,"Customer" = "casual"))

#diatas tidak perlu karena emg dr awal gitu
```

# Add columns that list the date, month, day, and year of each ride
# This will allow us to aggregate ride data for each month, day, or year ... before completing these operations we could only aggregate at the ride level
# https://www.statmethods.net/input/dates.html more on date formats in R found at that link

```{r}
all_trips$date <- as.Date(all_trips$started_at) #The default format is yyyy-mm-dd
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")

```

```{r}
colnames(all_trips)
```
# Add a "ride_length" calculation to all_trips (in seconds)
# https://stat.ethz.ch/R-manual/R-devel/library/base/html/difftime.html
```{r}
all_trips$ride_length <- difftime(all_trips$ended_at,all_trips$started_at)
```

```{r}
#Then the ride distance traveled in km 
all_trips$ride_distance <- distGeo(matrix(c(all_trips$start_lng, all_trips$start_lat), ncol = 2), matrix(c(all_trips$end_lng, all_trips$end_lat), ncol = 2))
all_trips$ride_distance <- all_trips$ride_distance/1000
```

# Inspect the structure of the columns
```{r}
str(all_trips)
```
```{r}
glimpse(all_trips)
```
```{r}
summary(all_trips)
```
```{r}
sum(all_trips$ride_length<0)
```

```{r}
#check missing value ride distance
sum(is.na(all_trips$ride_distance))
```

```{r}
sum(is.na(all_trips))
```

```{r}
all_trips$ride_length <- as.numeric(as.character(all_trips$ride_length))
is.numeric(all_trips$ride_length)

```


```{r}
summary(all_trips$ride_length)
```

```{r}
#cleaning data
all_trips_v2 <- subset(all_trips,ride_length>0 & !(is.na(all_trips$ride_distance)))
```

```{r}
sum(is.na(all_trips_v2$ride_distance))
```

```{r}
head(all_trips_v2)
```


# STEP 4: CONDUCT DESCRIPTIVE ANALYSIS
#=====================================
# Descriptive analysis on ride_length (all figures in seconds)

```{r}
mean(all_trips_v2$ride_length) #straight average (total ride length / rides)
```

```{r}
median(all_trips_v2$ride_length) #midpoint number in the ascending array of ride lengths
```

```{r}
max(all_trips_v2$ride_length) #longest ride
```
```{r}
min(all_trips_v2$ride_length) #shortest ride
```

```{r}
summary(all_trips_v2$ride_length)
```

# Compare members and casual users
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = mean)
```

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = median)
```

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = max)

```

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = min)
```

# See the average ride time by each day for members vs casual users
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)

```

# Notice that the days of the week are out of order. Let's fix that.
```{r}
all_trips_v2$day_of_week <- ordered(all_trips_v2$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

```

```{r}
head(all_trips_v2$day_of_week)
```

# Now, let's run the average ride time by each day for members vs casual users
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)
```

# analyze ridership data by type and weekday
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>%  #creates weekday field using wday()
  group_by(member_casual, weekday) %>%  #groups by usertype and weekday
  summarise(number_of_rides = n()							#calculates the number of rides and average duration 
  ,average_duration = mean(ride_length)) %>% 		# calculates the average duration
  arrange(member_casual, weekday)		
```

# Let's visualize the number of rides by rider type
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")

```

```{r}
#Fist we calculate the average distance, distance for both the casual and member type users:

userType_means <- all_trips_v2 %>% group_by(member_casual) %>% summarise(mean_duration = mean(ride_length),mean_distance = mean(ride_distance))

membervsduration <- ggplot(userType_means) + 
                   geom_col(mapping=aes(x=member_casual,y=mean_duration,fill=member_casual), show.legend = FALSE)+
                   labs(title = "Average travel duration by User type",x="User Type",y="Mean distance In seconds")
membervsdistance <- ggplot(userType_means) + 
                    geom_col(mapping=aes(x=member_casual,y=mean_distance,fill=member_casual), show.legend = FALSE)+
                    labs(title = "Average travel distance by User type",x="User Type",y="Mean distance In Km")

grid.arrange(membervsduration, membervsdistance, ncol = 2)
```


# Let's create a visualization for average duration
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
```

#=================================================
# STEP 5: EXPORT SUMMARY FILE FOR FURTHER ANALYSIS
#=================================================
# Create a csv file that we will visualize in Excel, Tableau, or my presentation software

```{r}
counts <- aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)
```

```{r}
counts
```
#=================================================
# STEP 5: EXPORT SUMMARY FILE FOR FURTHER ANALYSIS
#=================================================
# Create a csv file that we will visualize in Excel, Tableau, or my presentation software
# N.B.: This file location is for a Mac. If you are working on a PC, change the file location accordingly (most likely "C:\Users\YOUR_USERNAME\Desktop\...") to export the data. You can read more here: https://datatofish.com/export-dataframe-to-csv-in-r/

```{r} 
write.csv(all_trips_v2, "E:/DOKUMEN/DAxGOOGLE/STUDI KASUS/Jalur 1 SK1/all_trips_v2.csv")
```

```{r}
write.csv(counts, "E:/DOKUMEN/DAxGOOGLE/STUDI KASUS/Jalur 1 SK1/avg_ride_length.csv")
```

```{r}
all_trips_v3 <- subset(all_trips, select = -c(ride_length<0, start_station_name, end_station_name, start_station_id, end_station_id))
```


```{r} 
write.csv(all_trips_v3, "E:/DOKUMEN/DAxGOOGLE/STUDI KASUS/Jalur 1 SK1/all_trips_v3.csv")
```



